---
title: "Lab 2"
author: "Christopher Loan"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
#Ignore this code chunk for now. 
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

Run the following code to 
  (a) install the {nlme} and {janitor} packages
  (b) load the packages along with the tidyverse, and
  (c) access and quickly prep some data (from the {nmle} package) for plotting.

```{r eval=FALSE}
# Note: You only need to run this next line one time to install these two packages.
install.packages(c("nlme", "janitor")) #(a)
```


```{r data-download}

library(nlme) # (b)
library(janitor) # (b)
library(tidyverse) # (b)

pd <- Oxboys %>% # (c)
  clean_names() %>% 
  mutate(subject = factor(subject),
         occasion = factor(occasion)) %>% 
  filter(subject == "10" | subject == "4" | subject == "7") %>% 
  tbl_df()
```

1. Reproduce the following two plots, using the *pd* data. You can use whatever theme you want (I used `theme_minimal()`), but all else should be the same. 


```{r plots1, message = FALSE, warning = FALSE}

# Put code below for Plot 1. Note that Plot 1 is a line plot, not a smooth.
# pd %>% filter(subject == c('10','4','7'))

plot1 <-
  pd %>% ggplot(mapping = aes(x = age, y = height)) +
  geom_line() + theme_minimal() + facet_wrap(~subject) + ggtitle("Plot 1")
plot1
```
\newpage
```{r}
# Put code below for Plot 2. Note that Plot 2 is a line plot also.

plot2 <-
  pd %>% ggplot(mapping = aes(x = age, y = height, color = subject)) +
  geom_line() + theme_minimal() + ggtitle("Plot 2")

plot2
```

\newpage

2. Use the *mtcars* dataset from base R to replicate the following plots. (Just type *mtcars* into the console to see the dataset).

```{r plots2}

# Put code below for Plot 3
plot3 <- 
  mtcars %>% ggplot(mapping = aes(x = drat, y = mpg)) + 
  geom_point() + ggtitle("Plot 3") + theme_minimal()
plot3
```
\newpage
```{r}
# Put code below for Plot 4
plot4 <- plot3 + geom_smooth() + ggtitle("Plot 4")
plot4

```
\newpage
```{r}
# Put code below for Plot 5
plot5 <- plot3 + geom_smooth(method = "lm") + facet_wrap(~vs) + ggtitle("Plot 5")
plot5
```
\newpage
```{r}
# Put code below for Plot 6. I have given you the first line of code to start.
plot6 <- 
  ggplot(mtcars, aes(drat, mpg, color = factor(cyl))) + geom_point() + geom_smooth(method = "lm", se = F) + ggtitle("Plot 6") + theme_minimal()

plot6
```
\newpage


3. Use the *Orange* dataset, also part of base R, to replicate the following plots.
   
```{r orange-plots}
# Put code below for Plot 7
plot7 <- Orange %>% 
  ggplot(mapping = aes(x = age, y = circumference, color = Tree)) + 
  geom_point() + geom_line()+ theme_minimal() + ggtitle("Plot 7")
plot7 
```
\newpage
```{r}
# Put code below for the last plot. See slide 51 from the w2p2 class for labels.

plot8 <- Orange %>% 
  ggplot(mapping = aes(x = age, y = circumference)) + 
  geom_point(aes(color = Tree)) + 
  geom_smooth(method = "lm", se = F, color = "darkgray") + 
  theme_minimal() + 
  labs(x = "Age of Tree (in days)",
       y = "Circumference of the Trunk (in mm)",
       title = "Orange Tree Growth",       
       subtitle = "Gray line displays a linear model fit to the data"
       )
plot8
```
