---
title: "Loan EDLD 651 Lab 3"
author: "Chris Loan"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
```

1. Run the following code to load the *tidyverse*.

```{r load-tidyverse}
library(tidyverse)
```

1. Run the following lines of code to view and read about the data we'll use 
for this lab. 

Comment out line 26 after you run it, but puting a `#` at the start.

```{r gsscat, eval = FALSE}
str(gss_cat)
#View(gss_cat)
#?gss_cat 
```

2. Use `ggplot` to create a histogram of `tvhours`. The plot should:
  + Include a small amount of transparency so gridlines are visible under the 
    histogram.
  + Be a different color than the default
  + Include meaningful axis labels
  + Have a title and a subtitle

```{r question-2}
p1 <- ggplot(data = gss_cat, aes(x = tvhours)) +
  geom_histogram(fill = "blue", color = "black", alpha = 0.1) + 
  labs(x = "Hours of TV", 
       title = "Counts of TV hours",
       subtitle = "made from gss_cat dataset")
p1
```

3. Create a ridgeline density plot of `age` by `partyid`. Do you notice any strong
  differences by party affiliation?

```{r question-3}
library(ggridges)

p2 <- ggplot(data = gss_cat, aes(x = age, y = partyid)) + ggridges::geom_density_ridges()
p2
```

    * With only visual inspection, it's difficult to tell (and I feel I'm not good at density ridges yet; i.e., I can't tell what's about distribution vs. magnitude).
    * Though, it appears that older people are more likely to endorse strong affiliation with a party, compared to younger. 

4. Modify the plot from Question 3 to produce separate ridgeline densities by `race`. Do you see any patterns?

```{r question-4}
p3 <- ggplot(data = gss_cat, aes(x = age, y = partyid)) + geom_density_ridges() + facet_wrap(~race)
p3
```

    * To me, I still see the same pattern, with more partisan endorsement as age increases.
    * This trend appears similar across age, but again, I'm not certain it is significant.

5. Replicate the plots from Questions 3 and 4, but this time displaying the relations with boxplots (i.e., one boxplot from Question 3 and one boxplot for Question 4 in the chunk below). Which do you think displays the relation more clearly? Why?
  
```{r label, options}
p4 <- ggplot(data = gss_cat, aes(x = age, y = partyid)) + geom_boxplot() 
p4
p5 <- ggplot(data = gss_cat, aes(x = age, y = partyid)) + geom_boxplot() + facet_wrap(~race) 
p5
```


    * I personally prefer the boxplots, but that's mainly because I'm more confident in interpretting them.
    * I think the density plots are nice for showing a distribution is normal (or follows another distribution).

6. Run the following lines of code to get a correlation table as a data frame.

```{r corr-mat-df}
# You might need to install the ISLR package using the line below. 
# If so, uncomment out the line, install the pacakge, then comment the line again. 
# Your document won't knit if you do not comment the line below.

# install.packages(ISLR) 
library(ISLR)
cor_tbl <- Hitters %>% 
  select_if(is.numeric) %>% 
  cor(use = "pairwise.complete.obs") %>% 
  as.data.frame() %>% 
  mutate(row = row.names(.)) %>% 
  pivot_longer(
    cols = -row,
    names_to = "col",
    values_to = "cor"
  ) %>% 
  tbl_df()
```

Use the correlation table to create a heatmap using `geom_tile`. Change the 
colors to one of the four viridis palettes. What does the heatmap help you see?

```{r, heatmap}
library(viridis)
ggplot(cor_tbl, aes(x = row, y = col, fill = cor)) + geom_tile() + scale_fill_viridis_c()
```

7. The code below will produce a barplot. 

```{r barplot1}
ggplot(gss_cat, aes(rincome)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Use the code below to create a new data set - `gss_cat_new` - that redefines the `rincome` variable so that *Not applicable* is displayed first, followed by *No answer*, *Don't know*, *Refused*, and then each income category from lowest to highest. Once you've redefined the factor variable, re-produce the plot above by running the `ggplot` code at the bottom of the chunk. The plot should have *Not applicable* closest to the y-axis, then *No answer*, *Don't know*, etc.

```{r redfine-factor, echo = FALSE}
gss_relevel <- rev(levels(gss_cat$rincome))

gss_relevel <- c("Not applicable",
                 "No answer",
                 "Don't know", 
                 "Refused", 
                 gss_relevel[2:13])

gss_cat_new <- gss_cat %>% 
  mutate(rincome = factor(rincome, levels = gss_relevel))

ggplot(gss_cat_new, aes(rincome)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```