---
title: "Lab 4"
author: "Christopher Loan"
date: "October 14, 2020"
output:
  "pdf_document"
---

```{r setup, include=FALSE}
 knitr::opts_chunk$set(message = FALSE, 
                       warning = FALSE)
```


Install and load the package *Lahman*, which will give you access to the dataset *Teams*

```{r teams_echo, echo = TRUE}

library(Lahman)
library(tidyverse)
library(janitor)
```
* Produce a subset of the data (as a new object) that has the following characteristics:
    + Only one team (your choice)
    + data from 1980 to present (or as present as the dataset gets)
    + Includes 5 columns: name, yearID, W, L, R, RA

(The variables above correspond to the team name, the year, wins, losses, runs scored, and runs allowed)

* Make sure you select a team that is currently still around, or it probably
won't be interesting (see a list of current at [http://www.espn.com/mlb/teams](http://www.espn.com/mlb/teams)).

* Create a new variable corresponding to the winning percentage for the team 
  you chose over time

$$w_{pct} = \frac{wins}{wins + losses}$$

* Order by winning percentage: Least to greatest
* Order by winning percentage: greatest to least
* Compute the mean and standard deviation of winning percentage
* With the full dataset
  + compute the average and standard deviation of winning percentage for each team. 
  + Order by highest winning percentage
* Use the full data to reproduce the plot below

```{r select_team}
subset1 <- Teams %>% filter(teamID == teamID[length(Teams$teamID)], yearID >= 1980) %>% select(yearID, W, L, R, RA)
subset1
```

```{r win_percentage}
subset1 <- subset1 %>% mutate(win_perc = W/(W+L))
subset1
```

```{r reordering}
low_to_high <- subset1 %>% arrange(win_perc)
low_to_high
high_to_low <- subset1 %>% arrange(desc(win_perc))
high_to_low
```

```{r mean_and_sd}
subset1 %>% summarize(avg_wins = mean(win_perc))
subset1 %>% summarize(sd_of_wins = sd(win_perc))
```

```{r full_mean}
Teams %>% group_by(teamID) %>% mutate(win_perc = W/(W+L)) %>% summarize(avg_wins = mean(win_perc))
```

```{r full_sd}
Teams %>% group_by(teamID) %>% mutate(win_perc = W/(W+L)) %>% summarize(sd_of_wins = sd(win_perc))
```

```{r full_desc}
Teams %>% group_by(teamID) %>% mutate(win_perc = W/(W+L)) %>% summarize(avg_wins = mean(win_perc)) %>% arrange(desc(avg_wins))
```

```{r plot}
# Please put the code for the plot in this chunk.
Teams %>% 
  filter(name == 'Detroit Tigers'| 
           name == 'New York Yankees'|
           name == 'San Diego Padres') %>% 
  mutate(win_perc = W/(W+L)) %>%  
  ggplot(mapping = aes(x = yearID, 
                       y = win_perc, 
                       color = name)) + 
  geom_line() + labs(color = 'Team', 
                     x = 'year_id', 
                     y = 'w_pct')
```
