---
title: "Loan EDLD 651 Lab 10"
author: "Christopher Loan"
date: "11/15/2020"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    highlight: kate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
library(tidyverse)
library(here)
library(rio)
```

# Importing the Data & Binding (#4)
```{r data}
fotr_dat <- import(here("data", "The_Fellowship_Of_The_Ring.csv"))
tt_dat <- import(here("data", "The_Two_Towers.csv"))
rotk_dat <- import(here("data", "The_Return_Of_The_king.csv"))
lotr_tidy <- fotr_dat %>% bind_rows(tt_dat) %>% bind_rows(rotk_dat)
lotr_tidy 
```

# Is the data tidy? No. Why not? (#5)
The data above is not tidy because the columns "Male" and "Female" should both be listed on their own row.

# Tidying the data (#6)

```{r cleaning}
lotr_tidy <- lotr_tidy %>% 
  pivot_longer(cols = c(Female, Male), 
               names_to = 'Gender',
               values_to = 'Word_Count')

lotr_tidy %>% export(here("data", "lotr_tidy.csv"))
```

# Tidying the data (pt. 2: from Male.csv & Female.csv; #7)

```{r}
male <- import(here('data', 'Male.csv'))
female <- import(here('data', 'Female.csv'))

joined_dat <- male %>% 
  bind_rows(female) 

joined_dat <- 
  joined_dat %>% 
  pivot_longer(cols = c(Elf, Hobbit, Man),
              names_to = "Race", 
              values_to = "Word_Count") #%>% 
  
joined_dat <- 
  joined_dat %>% 
  relocate(c("Film", "Race")) %>% 
  arrange(desc(Word_Count), Race, Gender)

joined_dat$Race <- 
  ifelse(joined_dat$Race == "Man", 
         "Human", 
         joined_dat$Race)
```


## Do these two tables match perfectly? (Extra code chunk)
```{r equivalence}
sum(
  joined_dat %>% 
  arrange(desc(Word_Count), Race, Gender) != 
    lotr_tidy %>% 
    arrange(desc(Word_Count), Race, Gender)
  ) == 0
```

## What were the total number of words spoken by each "Race" of humanoid in the LOTR trilogy? 

```{r visualization, results = "asis"}
lotr_tidy %>% 
  group_by(Race) %>% 
  summarize(total_words = mean(Word_Count)) %>% 
  tibble() %>% knitr::kable(col.names = c("Race", "Total Words"))

lotr_tidy %>% 
  group_by(Race) %>% 
  summarize(total_words = mean(Word_Count)) %>% 
  ggplot(aes(x = Race, y = total_words, fill = Race)) + 
  geom_col() + 
  theme_minimal() + 
  labs(y = "Total Words Across Trilogy", 
       title = "Words Spoken by Race in LOTR Trilogy")
```

# Messying the data (#8)

```{r messying}
messy_dat1 <- lotr_tidy %>% 
  pivot_wider(names_from = Race, 
              values_from = Word_Count)
messy_dat1
messy_dat2 <- lotr_tidy %>% 
  pivot_wider(names_from = Gender, 
              values_from = Word_Count)
messy_dat2
```