---
title: "Loan EDLD 651 Lab 6"
author: "Christopher Loan"
date: "10/23/2020"
output: 
  pdf_document:
    highlight: espresso
---

```{r setup, include = F}
knitr::opts_chunk$set(
	echo = FALSE,
	fig.height = 4.5,
	fig.width = 6.5,
	message = FALSE,
	warning = FALSE,
	include = TRUE
)
```

```{r libraries}
# 2. Load Libraries
library(tidyverse)
library(here)
library(rio)
library(janitor)
theme_set(theme_minimal())
```

```{r data}
eclsk <- import(here('data', 'ecls-k_samp.sav')) %>% clean_names()
```

```{r plot, echo = F, include = T}
eclsk %>%
  filter(ethnic < 5) %>%
  arrange(desc(ethnic)) %>% 
  characterize() %>% 
  ggplot(aes(x = t1mscale, y = t2mscale)) +
  geom_point() + 
  geom_smooth(method = 'lm', aes(color = school_type))+
  facet_wrap(~ethnic) 

```

```{r descriptives}
t1mean <- mean(eclsk$t1mscale, na.rm = TRUE) 
t1sd <- sd(eclsk$t1mscale, na.rm = TRUE)
t2mean <- mean(eclsk$t2mscale, na.rm = TRUE) 
t2sd <- sd(eclsk$t2mscale, na.rm = TRUE)

average_gain <- t2mean - t1mean

```

The mean difference observed was `r average_gain`.

# Loading Data

Loading data is particularly easy when you use both the `{here}` and the `{rio}` packages. 

`here::here()` provides a heuristic to help you find the filepath needed for your data.

The code below will return simply the filepath for your current working directory:

```{r echo = T, eval = F}
here::here()
```

If you keep your data in a subfolder of your current working directory (e.g., let's call ours "data"), then you can reference that folder by including it in the function call:

```{r echo = T, eval = F}
here::here('data')
```

We're almost there; now we just need to give it the file name and import it. 

Similar to folder names, we specify this by putting the name of the document in quotes. Just separate the two with a comma. 

To actually import the data, we'll have to use `rio::import()`, which is a wrapper function for several importing functions from various packages.

Place this function around the `here::here()` call & don't forget to store the object with the `<-` operator!
```{r  echo = T, eval = F}
dat <- rio::import(here::here('data', 'datafile.csv'))
```